---
title: "Activity Data Treatment with R (30/03/2024)"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

## Activity to treat and load data in R

### Activity A

-   Create a DataFrame from the following vectors:

```{r}
X <- c(1,2,3,1,4,5,2)
Y <- c(0,3,2,0,5,9,3)
```

-   Installing package dplyr

```{r}
#install.packages("dplyr")
```
```{r}
dfA <- data.frame(X,Y)
dfA
```
-   Delete duplicates using an adequate R function `help(unique)`
```{r}
#We use unique for this part.
dfA <- unique(dfA)
dfA
```

-   Introduce a name for the rows. They will be: "CA","SE","MA","BA","VA"
```{r}
rownames(dfA) <- c("CA","SE","MA","BA","VA")
dfA
```
-   Create a new Z column which is the sum of the first two columns divided by the first one.
```{r}
dfA["Z"] <- (dfA["X"] + dfA["Y"])/dfA["X"]
#dfA$z <- NULL
dfA
```
-   Change the X value from province BA by 2
```{r}
dfA["BA","X"] <- 2
dfA
```
-   Select regions which Y value is less than 4
```{r}
dfA[dfA$Y < 4,]
```

-   Create a factor that says if it is a city from the coast or not (C or I) and add it as a column.
```{r}
costa <- c("C","I","C","I","C")
dfA["coast"] <- costa
dfA
```

### Activity B

-   Load the dataset from the URL <http://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data> having in consideration that missing values are "?" `url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data"`. If the URL do not work, do it as a local file called `imports-85.csv`.
```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data"
dfB <- read.table(url,header = FALSE, sep = ",", na.strings = "?")
dfB
```

-   Make a summary of the DataFrame.
```{r}
summary(dfB)
```

-   Which are the DataFrame dimensions?
```{r}
dim(dfB)
```
-   Inspect the data, which type is each class? Extract the distinct values that takes each factor.
```{r}
sapply(dfB, class)
```
-   Obtain the proportion of missing values by column using sum and is.na and an apply(with nrow).
```{r}
apply(apply(dfB,2,is.na),2,sum)
```
-   Modify the first column moving it to the 0-5 range.
```{r}
dfB["V1"] <- dfB["V1"] + 2
dfB["V1"]
```
-   Give a mean of the values from the first column compared to the number of doors of the car.
```{r}
tapply(dfB$V1, dfB$V6, mean)
```
-   Get the Audi cars which have four doors and their V13 column is less than 55.
```{r}
library(dplyr)
filter(dfB, dfB$V3 =='audi' & dfB$V6 =='four' & dfB$V13 < 55)
```
