---
title: "Inferencia estadística - Contrastes de hipótesis y poder estadístico"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
  html_document: default
---

# Applying different hypothesis tests

## Hypothesis contrast for sample means

### Contrast over sample mean

We will need to use $\mathit{t}$-Student for this case.

We have a dataset about lizards:

```{r}
lizard <- c(6.2, 6.6, 7.1, 7.4, 7.6, 7.9,
            8, 8.3, 8.4, 8.5, 8.6,
           + 8.8, 8.8, 9.1, 9.2, 
           9.4, 9.4, 9.7, 9.9, 10.2, 10.4, 10.8,
           + 11.3, 11.9)
hist(lizard)
```

Empleando el test t de Student con el siguiente comando obtenemos un p-valor muy pequeño. En consecuencia la hipótesis nula de que la media es 0 queda rechazada. Además nos da un intervalo de confianza al 95 sobre el parámetro de la media muestral.

The hypothesis contrast base case is that mean is 0, due to the base on "Normal" distribution. So, the hypothesis are formalized as:

-   Null Hypothesis or $H_{0}$: $\mu$ = 0.
-   Alternative Hypothesis or $H_{1}$: $\mu$ \<\> 0.
-   Level of significance $\alpha$ = 0.05.

We need to use the function "t.test()" which will give us a low p-value. Consequently, $H_{0}$ is rejected as our $\mathit{p}$-value is lower than $\alpha$, then we hae a 95% of statistical significance that our mean is not 0 *in our sample*.

```{r}
t.test(lizard)
```

Notice that for more confidence in our hypothesis we would need to change the value of $\alpha$ to a bigger value.

We can modify that level:

```{r}
t.test(lizard, conf.level = 0.99) # Increase to 99%
```

As the $\mathit{p}$-value is lower than 0.01, we reject definitely $H_{0}$.

But, we can define our own mean using the parameter "mu" in the "t.test()" function and do it in a lateral or bilateral mode:

```{r}
t.test(lizard, mu=8.4,conf.level = 0.9) #bilateral with mean 8.4,90% and H1: mu!=8.4 -> mu = 8.4
t.test(lizard, mu=8.4,alternative = "g" ,conf.level = 0.9) #lateral, H1: mu>8.4 -> mu <= 8.4
t.test(lizard, mu=8.4,alternative = "l" ,conf.level = 0.9) #lateral, H1: mu<8.4 -> mu >= 8.4
```

In these cases the $\mathit{p}$-values are bigger than 0.05 so $H_{0}$ remains, which means that mean is:

-   Equal to 8.4
-   More or equal than 8.4
-   Less or equal than 8.4

Lets see the atributes of `t.test`:

```{r}
a <- t.test(lizard, mu=8.4,conf.level = 0.9) 
a$conf.int
a$p.value
a$statistic
```

### Hypothesis contrast for mean´s difference between 2 normal samples.

In this case we will do a contrast between means in 2 normal samples.

-   $H_{0}: \mu_x - \mu_y \>= 0$
-   $H_{1}: \mu_x - \mu_y < 0$

In the first case, we will be supposing that we have different variances, so we need to do a Welch´s test.

```{r}
x <- rnorm(200,3,2)
y <- rnorm(200,4,1)
t.test(x,y,alternative = "l")   # Lower
```

As the $\mathit{p}$-value is really small, we accept $H_{1}$.

In the second case we suppose normal sample, independent and with the same variance. Is because of this that we can use the $\mathit{t}$-test for two samples.

```{r}
x <- rnorm(200,3,2)
y <- rnorm(200,2,2)
t.test(x,y,alternative = "l",var.equal = TRUE )
```

As the $\mathit{p}$-value is 1 and is bigger than $\alpha$, we accept $H_{0}: mu_x - mu_y \>= 0$.

## Statistical power analysis

The statistical power analysis relates the experiment to the statistical power by means of two lines of study.

A power analysis using Student's t-distribution is shown below. This technique requires normality in the samples and that they are independent.

-   The first question is, given some statistical characteristics such as $\alpha$, $\pi$ or the mean of the samples to be considered, what should be the sample size N of these samples to obtain these values?

Suppose we want to perform a study with laboratory mice in which we want to investigate the relationship between feeding. We prepare two types A and B, and the weight of the mouse. After performing our study, we expect that the mean weight of the mice with feeding A will be 25 grams, while with feeding B it will be 22 grams. We also want a significance of 0.05 and a statistical power of 0.80.

We will perform a power analysis based on Student's t-distribution that will tell us the sample size of each of the samples.

As a result of the analysis, we see that N = 3, so the experiment should have 3 mice taking type A feeding and 3 mice taking type B feeding.

```{r}
power.t.test(delta = (25-22), # diferencia entre las medias esperadas
             power = 0.80, # poder estadístico deseado
             sig.level = 0.05, # nivel de significación deseado
             type = "two.sample", # tipo de test
             alternative = "two.sided" 
             )
```

-   The second line of study is the reverse question, given statistical characteristics, not including $\pi$, and sample size N, **what statistical power will the experiment have**?

Continuing with the example of the mice, let us assume that the experiment will be performed with a sample of 5 mice per sample (10 mice in total), keeping the other parameters specified above except the statistical power at the same values.

Next, we will perform the statistical power analysis to show us the expected statistical power in this experiment.

As a result of the analysis, we see that the expected statistical power is 0.98.

```{r}
power.t.test(n = 5, # número de observaciones por grupo
             delta = (25-22), # diferencia entre las medias esperadas
             sig.level = 0.05, # nivel de significación deseado
             type = "two.sample", # tipo de test
             alternative = "two.sided" 
             )
```
